---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Chelsea Nelson"
date: today
format:
  html:
    
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
from: markdown+emoji 
editor: 
  markdown: 
    wrap: 72
---

::: {.callout-tip icon="false"}
## Github Repo Link

[Final Project 1
Repo](https://github.com/stat301-1-2023-fall/final-project-1-chelseanelson)
:::

```{r}
#| label: loading_packages 

library(tidyverse)
library(readxl)

```

## Data source

I will be using the Economic Policy Institute's data on family budget,
which also then corresponding to telling us about the cost of living in
each county in America. This dataset provides insights into the average
economical weights of different aspects of life for each county in
America both annually and monthly, whiles dividing 
these averages further by also looking at the
different family types as well. I have provided an additional copy of
the same dataset, with less columns and information from
Kaggle. This in case that the analysis that in which I wish to conduct starts to not follow the overall larger EPI dataset. Thus turning towards the smaller
one from Kaggle where I can focus in more on each variable, looking only at the annual costs and not both monthly and annually as in the larger EPI 
provided dataset. I additionally wanted to join this dataset 
with others that provide information on the minimum wage in each state 
and perhaps the racial majorities of each county, but I was unable to 
find suitable datasets, thus I will be adding these variables on myself during 
the first stages of my EDA. [^1]

[^1]: These datasets were sourced from [EPI Family Budget
    Map](https://www.epi.org/resources/budget/budget-map/) & [Kaggle US
    Cost of Living
    Dataset](https://www.kaggle.com/datasets/asaniczka/us-cost-of-living-dataset-3171-counties/data).

## Why this data

The reason I choose this dataset is because I thought it would be
interesting to see how the cost of living differs not only on the level
of state vs state, but looking further into the issue by seeing how cost
of living differs by family size and county location. Additionally, I
think this would be an opportunity to learn how the incomes levels and
expenses in each county and state differ from each other, while also
starting to understand why these differences are present. I further
think that by looking additionally at minimum wage in each state, it
will bring attention to a lot of the inequity present in the states.
Thinking on this now, I could also create a variable which states what
the majority race or ethnicity in each county is as well, I think this
would be an interesting statistical insight to find and correlate to
expense levels in the different counties and states.

## Data quality & complexity check

```{r}
#| label: loading_datasets

# larger version of dataset
fbc_data <- read_excel("data/raw/fbc_data_2022.xlsx", sheet = "County") %>% janitor::clean_names()
fbc_data_codebook <- read_excel("data/raw/fbc_data_2022.xlsx", 
                                sheet = "Codebook")
fbc_data
fbc_data_codebook

# smaller version of dataset --> only shows annual cost
cost_of_living_us <- read_csv("data/raw/cost_of_living_us.csv") 
cost_of_living_us

```

```{r}
#| label: skimming_the_data

fbc_data %>% skimr::skim()
fbc_data %>% naniar::miss_var_summary()

cost_of_living_us %>% skimr::skim()
cost_of_living_us %>% naniar::miss_var_summary()
```


The fbc_data there are 27 variables and 31,430 observations. Within this, there
are twenty-three numerical variables and four categorical variables. However,
I want to change the metro variable from being a numerical variable with only
0 and 1s to that of a logical later on. Five variables, which all pertain
to different rankings and extra information, share the same amount of missing 
data. However, I feel that since it is such a small percentage, with each 
missing only 10 values (0.3181) respectively  of their portion of the data, 
I will see how I can adapt or change my analysis to either ignore or get
rid of this missingness altogether.

Looking at the smaller version of the data, which looks only at the annual 
costs has 15 variables and 31430 observations. Within this, there are five
categorical variables and ten numerical variables. There is only one variable
which has missing values, being the median_family_income variable. It is 
missing ten instances (0,3181 of the data), thus since it is a small number 
I will see how I can operate around this missingness. 
This variable was also one of the five missing in the larger dataset, thus this helps to prove that this dataset is the same as the larger fbc_data dataset, 
but just showcasing specific variables. 

## Potential data issues

I think the main issues that I will face is being able to figure out how 
to navigate looking at all the different levels that my data is assessing, 
relating to the different family types, different counties, different states, 
and how they differ between annual and monthly cost. 

Additionally, I feel that I might struggle slightly with wanting to add in my 
own variables of minimum wage and the racial composition or racial majority of 
each county into my dataset and analysis. 

Furthermore, I have to look more into where the missing data for the variables
lie, and how I can navigate around them in order to still provide similar and
consistent comparisons for each county and family type. 

## Misc

As I have been discussing above, I plan to add other information such as 
minimum wage for each state and the ethnic makeup or majority in each county,
if I am able to find this for each or most counties in America. I plan to add
the minimum wage through using the case when function, however I am unsure
of how I want to add the ethnic makeup variable to my dataset. Currently,
I think I will make my own dataset in excel, and then join it together with 
my current dataset. 

Furthermore, I currently feel that I want to look at how 
minimum wage of each state corresponds with the annual cost in each county and 
how being in a metro area makes this either higher or lower. 

Thinking about this, I was wondering if I should look to see if different 
metro areas have different minimum wages than the state as a whole as well. 




