---
title: "Cost of Living Throughout America"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Chelsea Nelson"
date: today
format:
  html:
    
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
from: markdown+emoji 
editor: 
  markdown: 
    wrap: 72
---
::: {.callout-tip icon="false" appearance="simple"}
## Github Repo Link

[Final Project 1
Repo](https://github.com/stat301-1-2023-fall/final-project-1-chelseanelson)
:::

```{r}
#| label: loading-packages 
library(tidyverse)
library(DT)
```

```{r}
#| label: loading-data 

fbc_data <- read_rds("data/fbc_data.rds")
# fbc_data_codebook <- read_csv("data/fbc_data_codebook.csv")
# still need to make the codebook
```

## Introduction

In a comprehensive exploration of the EPI Family Budget dataset, this report delves into the intricate dynamics of cost of living variations across geographical regions, shedding light on the nuanced relationships between family budgets, income disparities, and metro classifications in the United States.

I was originally motivated to perform this analysis, as I think it is interesting and beneficial to understand and see how the cost of living differs not only on the level of state vs state, but also looking further into the issue by 
seeing how cost of living differs by family size and county location. Additionally, I think this would be an opportunity to learn how incomes levels 
and expenses in each county and state differ from each other,starting to understand why these differences are present. I further think that by looking additionally at minimum wage in each state, I think that 
the analysis I conducted will bring attention to a lot of the inequity 
present in the United States of America. 

In terms of initial curiosities while conducting early research on this data,
I was interested in looking at how a median family income for each county then correlates to the total annual expenses on the state, regional, and metro levels.
Furthermore, to then see if there are any patterns or trends in budget allocation
that stand out or are different from others on multiple levels, such as the county, metro, state, and regional levels. Moving from looking at annual to monthly numbers, I was interested in seeing if there is any change between the 
two calculations, and if so, how that changes the overall cost of living in the different areas of interest that I have presently stated. Through having these
starting curiosities, I was able to conduct a full exploration of the data that
focused on comparisons and correlations to help bring insight into how different
categories of expenses are valued and allocated in relation to the total
expenses and cost of living in different geographical locations. 

As stated above I will be using the Economic Policy Institute's data on family budget, which also then corresponding to telling us about the cost of living in
each county in America. This dataset provides insights into the average
economical weights and costs of different aspects of life for each county in
America both annually and monthly, whiles dividing these averages further by 
also looking at the different family types as well, ranging from 1 parent and no children families to 2 parent 4 children families. In order to enhance the 
dataset for my own research, I added information on the geographical region in
which each county is located based on their state 
(south, midwest, northeast, and west) and added the minimum wage of each 
pair of state and county. The former information was sourced from the USA 
Census Website, and the latter was sourced from Paycom.com. See [References @sec-references] for additional information and citation on the Economic Policy Institute's dataset, as well as the extra information I obtain for the addition of my region and minimum wage variables. 

In terms of the layout of my report, I will first discuss and provide an 
overview and quality check of my data, being descriptive of how the data looks
and how I formatted it in the best way for my own research. I will then start
my main explorations, where I will discuss on early univariate and bivariate a
analyses in which I conducted, and then turn the attention towards three separate main questions that constructed and established the flow of my exploratory data analysis. Lastly, I will conclude with a summary of the main insights that I 
have founded throughout my research, as well as discussing potential directions 
that the analyses in which I conducted can be taken to the next level. 

## Data Overview & Quality 

The FBI Family Budget dataset in its original state consisted of 27 variables 
and 31,430 observations. Within this, there were twenty-three numerical variables and four categorical variables. However, I did add my own minimum wage and regional variables, as well as the changed the variable type of the family type and metropolitan status variables. Additionally, I made sure that I tackled 
how I was going to work around the missingness that was in my dataset. Therefore, after further investigation, I realized that all of the missing values for my variables corresponded to one specific county and its multiple different family cases. Thus in this case, I decided it would be best to fully remove the observations of that particular county from my dataset, as I felt leaving it in would case more problems in terms of furthering my analysis than taking it
out. Thus my updated version of the dataset includes 29 variables with 
31,420 observations. Within this, there are six categorical variables and 
twenty-three numerical variables. Thus after my manipulation, the dataset is of high quality, being extremely well-made and will be easy to use during my 
data analyses as there are no underlying issues or problems. 

## Explorations 

This is the heart or main body of the the report. The structure of this portion can look VERY different depending on narrative choices and organization choices made to guide readers through the project EDA. Could decide to make sections for major parts of the EDA and not used an Explorations header. The key is that you have sections that are clearly organized and guide the reader through a systematic and thoughtful exploration of the dataset(s).

A thoughtful exploration weaves together prose, figures, and tables to help a reader build understanding. It is a guided tour that should have clear motivations for each step in the process (think guiding curiosities or research questions). It should be clear how things build of of each other and why we are looking at what is being presented. Should also be letting the reader know whether what is shown is as expected or is it surprising (maybe some of each).

You could imagine structuring this section around analysis type: univariate, bivariate, and multivariate. Can be a little harder to build a narrative with this structure. Could build sections that are dedicated to important variables: variable 1, variable 2, etc. In this case the sections would thoroughly explore the section’s variable through univariate, bivariate, and multivariate analyses. Alternatively, could build your sections around broad curiosities/research questions. Some of which probably inspired the project, and others that were developed through conducting the EDA.

## Conclusions 

State conclusions or insights. Were you surprised by things you found or were they as expected? Why? This is a great place for future work, new research questions, and next steps.

# fix my connection here
## References {.unnumbered #sec-references}
  1. Economic Policy Institute (2022, March) Family Budget Map. [https://www.epi.org/resources/budget/budget-map/](https://www.epi.org/resources/budget/budget-map/)

  2. U.S. Census Bureau (2021, October 8). Census Regions and Divisions of the United States. 
[https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf)

  3. Paycom (2023, October 2). Your 2023 Guide to Every State's Minimum Wage. [https://www.paycom.com/resources/blog/minimum-wage-rate-by-state/](https://www.paycom.com/resources/blog/minimum-wage-rate-by-state/) 

# will probably get rid of this 
## Appendix: Technical Info

A place to share complex and important technical steps that may highly impact explorations, but the details are too technical to share in main body of the report.

### Appendix: Extra Explorations 

Useful in cases where there are an over-abundance of explorations and they are not useful in the main body of the report or they are uninteresting, but still think readers should have access to them for reference.
