---
title: "Cost of Living Throughout America"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Chelsea Nelson"
date: today
format:
  html:
    
    toc: true
    embed-resources: true
execute:
  echo: false
  warning: false
from: markdown+emoji 
editor: 
  markdown: 
    wrap: 72
---
::: {.callout-tip icon="false" appearance="simple"}
## Github Repo Link

[Final Project 1
Repo](https://github.com/stat301-1-2023-fall/final-project-1-chelseanelson)
:::

```{r}
#| label: loading-packages 
library(tidyverse)
library(DT)
```

```{r}
#| label: loading-data 

fbc_data <- read_rds("data/fbc_data.rds")
# fbc_data_codebook <- read_csv("data/fbc_data_codebook.csv")
# still need to make the codebook
```

## Introduction

In a comprehensive exploration of the EPI Family Budget dataset, this report delves into the intricate dynamics of cost of living variations across geographical regions, shedding light on the nuanced relationships between family budgets, income disparities, and metro classifications in the United States.

I was originally motivated to perform this analysis, as I think it is interesting and beneficial to understand and see how the cost of living differs not only on the level of state vs state, but also looking further into the issue by 
seeing how cost of living differs by family size and county location. Additionally, I think this would be an opportunity to learn how incomes levels 
and expenses in each county and state differ from each other,starting to understand why these differences are present. I further think that by looking additionally at minimum wage in each state, I think that 
the analysis I conducted will bring attention to a lot of the inequity 
present in the United States of America. 

In terms of initial curiosities while conducting early research on this data,
I was interested in looking at how a median family income for each county then correlates to the total annual expenses on the state, regional, and metro levels.
Furthermore, to then see if there are any patterns or trends in budget allocation
that stand out or are different from others on multiple levels, such as the county, metro, state, and regional levels. Moving from looking at annual to monthly numbers, I was interested in seeing if there is any change between the 
two calculations, and if so, how that changes the overall cost of living in the different areas of interest that I have presently stated. Through having these
starting curiosities, I was able to conduct a full exploration of the data that
focused on comparisons and correlations to help bring insight into how different
categories of expenses are valued and allocated in relation to the total
expenses and cost of living in different geographical locations. 

As stated above I will be using the Economic Policy Institute's data on family budget, which also then corresponding to telling us about the cost of living in
each county in America. This dataset provides insights into the average
economical weights and costs of different aspects of life for each county in
America both annually and monthly, whiles dividing these averages further by 
also looking at the different family types as well, ranging from 1 parent and no children families to 2 parent 4 children families. In order to enhance the 
dataset for my own research, I added information on the geographical region in
which each county is located based on their state 
(south, midwest, northeast, and west) and added the minimum wage of each 
pair of state and county. The former information was sourced from the USA 
Census Website, and the latter was sourced from Paycom.com. See [References @sec-references] for additional information and citation on the Economic Policy Institute's dataset, as well as the extra information I obtain for the addition of my region and minimum wage variables. 

In terms of the layout of my report, I will first discuss and provide an 
overview and quality check of my data, being descriptive of how the data looks
and how I formatted it in the best way for my own research. I will then start
my main explorations, where I will discuss on early univariate and bivariate a
analyses in which I conducted, and then turn the attention towards three separate main questions that constructed and established the flow of my exploratory data analysis. Lastly, I will conclude with a summary of the main insights that I 
have founded throughout my research, as well as discussing potential directions 
that the analyses in which I conducted can be taken to the next level. 

## Data Overview & Quality 

The FBI Family Budget dataset in its original state consisted of 27 variables 
and 31,430 observations. Within this, there were twenty-three numerical variables and four categorical variables. However, I did add my own minimum wage and regional variables, as well as the changed the variable type of the family type and metropolitan status variables. Additionally, I made sure that I tackled 
how I was going to work around the missingness that was in my dataset. Therefore, after further investigation, I realized that all of the missing values for my variables corresponded to one specific county and its multiple different family cases. Thus in this case, I decided it would be best to fully remove the observations of that particular county from my dataset, as I felt leaving it in would case more problems in terms of furthering my analysis than taking it
out. Thus my updated version of the dataset includes 29 variables with 
31,420 observations. Within this, there are six categorical variables and 
twenty-three numerical variables. Thus after my manipulation, the dataset is of high quality, being extremely well-made and will be easy to use during my 
data analyses as there are no underlying issues or problems. 

## Explorations 

Welcome to the heart of my analysis – a comprehensive exploration of the EPI Family Budget dataset. In this section, we embark on a journey through the intricate layers of data, unraveling the complexities of cost-of-living variations across multiple different geographical facets in the United States.

Before diving into our main questions, let's briefly revisit the insights gleaned from our preliminary analyses. Unraveling the individual and paired variables 
that have provided us with a foundational understanding of the dataset's landscape. 

### Univariate Analysis


### Bivariate Analysis 


Now, with this groundwork established, we can turn our attention to the core questions that drive my exploration:

### 1. How does the cost of living vary across different geographical facets and metro classifications? 

For myself: first look at the largest level so national, then regional, then the
states in every region comparing them. followup question: Are there observable trends in transportation costs based on the availability and accessibility of public transportation in metro areas?

Our first inquiry delves understanding how the geographical tapestry and metro classifications of different areas in America help to uncover the intricate variations in the cost of living. Thus through this question, I aim to unravel
the economic nuances that define household budgets. 

To start my investigation for this question


Say an intro sentence here:

### 2. Are there discernible patterns or trends in budget allocation that stand
out, considering family types and monthly expenses? 


Building on the foundation of my previous analyses, I now shift focus more heavily on the allocation of expenses. Exploring patterns and trends in budget allocation, I aim to identify distinctive markers that stand out amidst the diverse financial landscape pertaining to the different family types and the 
relationship between monthly and annual expenses. Thus, looking to see how 
family size impacts budget allocation. And if there are specific categories 
where larger families allocate a significantly higher percentage of their budget.

For me: I will work on the heatmap and see how that looks. Again look from the 
national, regional, and state levels. And then also I want to pick two states one
being the best in a category and one being the worst and then look at how the counties compare for those two states as well. Follow-up question: What is the impact of the number of working adults in a family on budget allocation? Do dual-income families allocate their budgets differently compared to single-income families?

intro statement:

### 3. How does the minimum wage in different states correlate with the affordability of living, particularly in terms of housing, healthcare, and other essential expenses for various family types and regions?

Lastly, I turned my focus to looking at a critical economic variable -- minimum wage. Through looking at minimum wages correlates to the cost of living, I aimed to shed light on how variations in minimum wage impact crucial aspects of family budgets, from housing and healthcare to other essential expenses. 


Thus through answering these three esstenial questions, I was able to gain interesting insights on the following.....

## Conclusions 

State conclusions or insights. Were you surprised by things you found or were they as expected? Why? This is a great place for future work, new research questions, and next steps.



## References {.unnumbered #sec-references}
fix my connection here

  1. Economic Policy Institute (2022, March) Family Budget Map. [https://www.epi.org/resources/budget/budget-map/](https://www.epi.org/resources/budget/budget-map/)

  2. U.S. Census Bureau (2021, October 8). Census Regions and Divisions of the United States. 
[https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf)

  3. Paycom (2023, October 2). Your 2023 Guide to Every State's Minimum Wage. [https://www.paycom.com/resources/blog/minimum-wage-rate-by-state/](https://www.paycom.com/resources/blog/minimum-wage-rate-by-state/) 

## Appendix: Extra Explorations 

Useful in cases where there are an over-abundance of explorations and they are not useful in the main body of the report or they are uninteresting, but still think readers should have access to them for reference.
